---
title: "Demographic Trends in U.S. Border Encounters"
subtitle: "A Temporal and Regional Analysis of CBP Data"
author: "Campbell Bowers"
institute: "Houghton University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
library(tidyverse)
library(janitor)

cbp_resp <- bind_rows(
  read_csv("https://www.cbp.gov/sites/default/files/assets/documents/2023-Nov/nationwide-encounters-fy20-fy23-aor.csv"),
  read_csv("https://www.cbp.gov/sites/default/files/2024-10/nationwide-encounters-fy21-fy24-aor.csv")
) |>
  janitor::clean_names() |>
  unique()
```

```{r, echo=FALSE}
style_xaringan(
  title_slide_background_image = "img/confetti.jpg"
)
```


class: center, middle

## Which demographic groups are most frequently encountered at the border? How have these encounters changed over time and by region? 

---

class: inverse, center, middle

# Dataset Description

---

class: inverse, center, middle

# Methodology

---

class: top, center

# Summary Statistics

```{r, echo=FALSE}
library(dplyr)
cbp_resp %>% 
  group_by(demographic) %>% 
  summarise(mean_encounters = mean(encounter_count, na.rm = TRUE),
            total_encounters = sum(encounter_count, na.rm = TRUE),
            count = n())
```

---
# Line Plot


```{r, echo=FALSE, mychunk, warning=FALSE, message=FALSE}
demographic_trends <- cbp_resp %>%
  group_by(fiscal_year, demographic) %>%
  summarize(total_encounters = sum(encounter_count, na.rm = TRUE)) %>%
  ungroup()

ggplot(demographic_trends, aes(x = fiscal_year, y = total_encounters, group = demographic, color = demographic)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Trends in CBP Encounters by Demographic Group",
       x = "Fiscal Year",
       y = "Total Encounters",
       color = "Demographic Group")
```

---
# Box Plot


```{r, echo = FALSE}
ggplot(cbp_resp, aes(x = land_border_region, y = encounter_count, fill = land_border_region)) +
  geom_boxplot() +
  labs(title = "Distribution of Encounters by Border Region",
       x = "Land Border Region",
       y = "Encounter Count") 
```

---

class: inverse, center, middle

# Limitations

---

class: inverse, center, middle

# Conclusion



